### 2023-11-28 methods.qmd
Testing data sources

---
title: "Methods"
author: "Jacqueline Grout"
format: html
editor: visual
---

## GP Survey

Technical annex:

Q55 base: Smoking habits - Total responses

Q55_1pct: Smoking habits - % Never smoked

Q55_2pct: Smoking habits - % Former smoker

Q55_3pct: Smoking habits - % Occasional smoker

Q55_4pct: Smoking habits - % Regular smoker

<https://gp-patient.co.uk/downloads/2023/GPPS_2023_Technical_Annex_PUBLIC.pdf>