### 2024-03-04 findings.qmd
Tweak rate charts

---
title: "Inequalities by ethnicity in the CHD pathway"
lang: en-GB
author: "Jacqueline Grout"
date: last-modified
date-format: "YYYY-MM-DD"
format:
  html:
    grid:
      sidebar-width: 250px
      body-width: 800px
      margin-width: 250px
      gutter-width: 1.5rem
    embed-resources: true
    smooth-scroll: true
    theme: cosmo
    fontcolor: black
    toc: true
    toc-location: left
    toc-title: Contents
    toc-depth: 3
editor: visual
execute:
  echo: false
  message: false
  warning: false
  freeze: auto
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false

library(targets)

library(tidyverse)
```

## Clustered GP Practices

```{r}
tar_read(cluster2_map)
```

### Cluster descriptions

#### Cluster 1 - Least diverse

This cluster is the least diverse of the five and the median percentage of White patients is 97% (94% White British).

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_1)
```

## Fourteen categories

94% White British

```{r}
tar_read(cluster2_eth_chart_1)
```
:::

#### Cluster 2

In this cluster the median percentage of White patients is 93% (87% White British). Those patients whose ethnicity is White are more likely to be White Irish or other White ethnicities (0.7% and 4% respectively) when compared to cluster 1. The median percentage of patients with a mixed ethnicity in this cluster is 2%.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_2)
```

## Fourteen categories

87% White British

```{r}
tar_read(cluster2_eth_chart_2)
```
:::

#### Cluster 3

In this cluster the median percentage of White patients is 78% (69% White British). The median percentage of patients with a mixed ethnicity in this cluster is 3.8%, which is higher than in cluster 2. Indian patients make up 3.3% of the patients, and 2.6% are patients whose ethnicity is Black African.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_3)
```

## Fourteen categories

69% White British

```{r}
tar_read(cluster2_eth_chart_3)
```
:::

#### Cluster 4

In cluster 4 just over half the patients are White with a median percentage of 55% (35% White British). The ethnic group of White has fewer White British than clusters 1 to 3, with 17% of patients having an ethnicity of White Other The median percentage of patients whose ethnicity is Black African in this cluster is 8.3%, which is higher than in cluster 3 and the median percentage of Black Caribbean patients is 3.9%. The median percentage of Black and Asian is 29% in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_4)
```

## Fourteen categories

35% White British

```{r}
tar_read(cluster2_eth_chart_4)
```
:::

#### Cluster 5 - Most diverse

This cluster is the most diverse. Overall the median percentage of Asian patients in this cluster is 45%. There are 13.3% Indian, 9.7% Pakistani and 2.9% Bangladeshi. Mixed and other ethnic groups form a large proportion of the patients in this cluster.

::: panel-tabset
## Five categories

```{r}
tar_read(cluster2_treemap_5)
```

## Fourteen categories

22% White British

```{r}
tar_read(cluster2_eth_chart_5)
```
:::

## Relative Index of Disparity

```{r}
tar_read(rel_iod_chart)
```

## Rate

### Risk factors

```{r}
tar_read(rate_chart_risk_fact)
```

### Risk factor identification

```{r}
tar_read(rate_chart_risk_fact_ident)
```

### Primary prevention

```{r}
tar_read(rate_chart_prim_prevent)

```

### Disease identification

```{r}
tar_read(rate_chart_disease_ident)
```

### Secondary prevention

```{r}
tar_read(rate_chart_second_prevent)
```

### Tertiary prevention

```{r}
tar_read(rate_chart_tert_prevent)
```

### Intermediate outcome

```{r}
tar_read(rate_chart_int_out)
```

### Full outcomes

```{r}
tar_read(rate_chart_full_out)
```