### 2024-04-25 deaths_explanation.qmd
Updated deaths notes

---
title: "deaths_explanation"
format: html
editor: visual
---

## Question

Why are the rates of CHD deaths and CHD hospital deaths lower in the 2 most diverse clusters, yet, "South Asians living in the UK have a higher rate of [coronary heart disease](https://www.bhf.org.uk/informationsupport/conditions/coronary-heart-disease "https://www.bhf.org.uk/informationsupport/conditions/coronary-heart-disease") and [diabetes](https://www.bhf.org.uk/informationsupport/risk-factors/diabetes "https://www.bhf.org.uk/informationsupport/risk-factors/diabetes") compared to White Europeans".

This study refers to the incidence rate.

## Project

The method of rate calculation is simply to take the amount of activity in the metric and divide this by the number of patients on the gp list expected to have CHD (calculated by multiplying the PHE synthetic prevalence estimate % by the GP list size).

The PHE synthetic prevalence calculations take ethnicity and age and sex into account in their methodology.

The factors controlled for in the prevalence model are:

-   age

-   sex

-   ethnicity

-   hypertension

-   diabetes

-   dyslipidemia

-   obesity

-   smoking

-   CKD

-   physical activity

-   family history

-   area deprivation (townsend score)

SW - Ethnicity is a covariate in the prevalence model although the variable only has two levels, white / non-white.  The effect is quite large.  They estimate that a non-white individual is 33% more likely to develop CHD than a white person with similar risk factors (age, sex, family history, etc etc).

Since their prevalence estimates account for the fact that a non-white patient would have 33% higher likelihood of developing CHD then this feeds into the denominator of our calculations and the larger denominator leads to a lower rate. This increased risk of developing CHD offsets the lower age profile of ethnic minority populatons (which is why the distribution of list size over the clusters is similar to the distribution of prevalent cases).  But we are still saying that having controlled for prevalence, the more diverse clusters have lower death rates.  We might have expected that more diverse clusters would have similar or higher death rates having controlled for prevalence.

We have calculated Age and Sex Standardised Mortality Rates (direct) for each of the clusters for CHD Deaths and All Cause deaths. These calculations used the same death data as was used in the numerator of the rates from the project.

SC - Prevalent population of the most diverse clusters is younger then would expect fewer deaths.

SC - Prevalence is a function of incidence and mortality and survival is a curve

SW - Are they dying of other causes? = No

![](secret/asmrchdprojrates.png){fig-align="center"}

JG - The table on the right shows the CHD deaths (same data that forms the numerator of the project calculation) as a crude rate per 10,000 population, as well as all cause deaths as crude rate. The table on the left is mortality rates age and sex standardised using census population. So, if the age and sex standardised mortality rates show the most diverse cluster to have the highest rate then why do the project rates using need not follow the same pattern?

SC - The standardised rate is saying if the populations have no difference in age and sex then cluster 5 is higher than you would expect, too many deaths for its age/sex profile. It is comparing to the total population whether people have the condition or not.

SC -The project chd rate is a case fatality rate, deaths in those that have the condition, this is saying that the highest observed deaths per person with the condition is in cluster 1. It's not comparing like with like and it's not age standardised so could be argued that cluster 1 could be higher because its prevalent population are older.

SC - higher incidence but better survival is possible

SC - \*\*the project rates are different to asmr because the asmr are population age standardised but the case fatality in the project that we are comparing to are not \*\* . It's complicated by not comparing like with like as the first is a measure of the rate death from CHD occurs in the population and the second is a measure of how many people already with the condition die.

On register:

![](secret/cvdreg.png){fig-align="center"}

Deaths:

![](secret/deaths.png){fig-align="center"}

Ecological Fallacy:

link to Robinson's paradox

ASMR only accounts for age and sex, it is partial and biased, doesnt account for need. CHD prevalence as a need denominator does control for other influencing factors such as deprivation, ethnicity etc.

Next steps

Take 2 practices and examine

Under 75 mortality and another indicator split it by deprivation and by clusters